---
title: 'Homework #3'
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dora et al. (European Spinal Journal (2002)) investigated spinal canal dimensions in subjects symptomatic with disc herniation selected for a discectomy and asymptomatic individuals (control group).  One of the areas of interest was determining if there is difference between the two groups in the spinal canal cross-sectional area (cm2) between vertebra L5/S1.  The data in the following table are simulated to be consistent with the results reported in the paper.  This data has been saved in the file entitled Hwk03.csv.  

a) Compute the mean, median, standard deviation, and range for the data in each group.
```{r}
# import data
hern <- c(2.62,2.57,1.98,3.21,3.59,1.60,1.80,3.91,
          2.56,1.53,2.39,2.67,3.53,2.26,2.82,2.09,
          3.79,2.45,2.55,2.10,2.28,2.33,2.81,3.70,
          2.61) 

```


```{r}

# calculate the mean, median, sd and range of the data
mean(hern)
median(hern)
sd(hern)
max(hern)-min(hern)
```
The mean of Herniated Disc Group (#1): 2.63

The median of of Herniated Disc Group (#1): 2.56

The standard deviation of Herniated Disc Group (#1): 0.666977

The range of Herniated Disc Group (#1): 2.38

```{r}
# import data
contr <- c(3.72,4.30,2.87,3.87,2.73,5.28,1.33,2.36,
           3.67,1.64,3.54,3.63,4.26,3.08,3.32,4.00,
           2.76,3.58,5.02,3.62,3.02,3.15,3.57,2.37,
           5.42,3.35,2.62,3.72,4.37,5.28,4.97,2.58,
           2.25,3.12,3.43,12.98,3.95,2.98,4.11,3.08,
           2.22)

#calculate mean, median, sd and range of the data
mean(contr)
median(contr)
sd(contr)
max(contr)-min(contr)

```
The mean for control group (#2): 3.685854

The median for control group (#2): 3.54

The standard deviation for control group (#2): 1.763087

The range for control group (#2): 11.65

b) Construct an appropriate boxplot for this dataset.
Boxplot for Herniated Disc Group (#1):
```{r}
# create a boxplot
boxplot(hern, col="orange", horizontal = TRUE)

```

Boxplot for Control Group (#2):
```{r}
# create a boxplot
boxplot(contr, col="orange", horizontal = TRUE)

```

c) Assess the normality of the data in each group and discuss your conclusions.

Normality of Herniated Disc Group (#1):
```{r}
# create a qq line to display the data for group 1

qqnorm(hern,ylab="Spinal Canal Dimensions",xlab="Scores",main="Herniated Disc Group (#1)") 
qqline(hern)

```

Conclusions on the normality of the data:
In order to have a normally distributed data, you have to see a straight diagonal line in the plot. In this case, most of the values fall in the line, but at the end we have 6 values that are in the left part of the line. We can say that is a left-skewed data. 

Normality of Control Group (#2):
```{r}
# create a qq line to display the data for group 1

qqnorm(contr,ylab="Spinal Canal Dimensions",xlab="Scores",main="Control Group (#2)") 
qqline(hern)

```

Conclusions on the normality of the data:
As mentioned before, in order to have a normally distributed data, you have to see a straight diagonal line in the plot. In this case, we have a similar plot as the Group 1. In this case, the plot follows a straight line, but it is going most of the points to the left of the plot. In addition, we have a value that is all the way to the left: 12.98. We have a left-skewed data. 


Knowing what you know from this class (so far) do you think there is a difference in the spinal canal cross-sectional areas between the population of subjects with disc herniations and the population of those who do not have disc herniations? 

In order to determine if there is a difference in the spinal canal cross-section ares between the population of subjects for both groups we have to conduct an Hypotesis test of 2 different independent samples.We will use the average of both groups, the standard deviation for both groups and the number of samples in both groups. 

Ho: Average of area in Group 1 and 2 are the same
Ha: Average of area in Group 1 and 2 are different

Test Statistic (Calculator): 

t = -3.45097
p = 0.00107

Since we have a 2-tail test we have to multiply p times 2

p = 0.00107*2 = 0.00214

Reject Ho since p-value is less than 0.05. 

Conclusion: There is sufficient evidence to conclude that there is a difference in the spinal canal cross-sectional areas between the population of subjects with disc herniations and the population of those who do not have disc herniations. 

