---
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# create the vectors with the data
gend <- matrix(c(1044,893,971,851,962,912,408,332,389,380,390,436), ncol = 6, byrow = TRUE)
colnames(gend) <- c("2020","2019","2018","2017","2016","2015")
rownames(gend) <- c("Female", "Male")

# make the matrix a table with as.table() function
gend <- as.table(gend)

# Display data fo gend
gend

```

```{r}
# Perform the chisq.test
results <- chisq.test(gend)

# display results
results$expected
results

```

Ho: No relationship exist between the number of male and female students over the last 5 years
Ha: There is a relationship between the number of male and female students over the last 5 years

Test Statistic = X-squared = 11.814
p-value = 0.03743

Conclusion: Reject Ho in favor of Ha. There is sufficient evidence to conclude that a relationship exist between the year and the amount of males and females.

```{r}
# create a barplot with the information

barplot(gend, ylim = c(0,1300),ylab='Counts',beside = TRUE, legend = TRUE)

```


